New lower bounds for Schur and weak Schur numbers

### 2021-06-25 article/weakschur.tex
Nouveau style biblio
2 typos (on se sent utile)

\begin{itemize}
\item Function \(g_{|[\![b + 1, a + b]\!]}\) is a sum-free colouring. Indeed, let \((x, y) \in 
	[\![b + 1, a + b]\!]^2\) such that \(g(x) = g(y)\). If \(g(x) = n + 1\) then \(z = x + y > 2 b + 1\) 
	and therefore, either \(z > a + b\) or \(f(z) \neq n + 1\). Otherwise, \(x + y > a + b\).
\item Function \(g\) is a weakly sum-free colouring. Indeed, let \((x, y) \in [\![1, a + b]\!]^2\) such that 
	\(z = x + y \leqslant a + b\) and \(g(x) = g(y)\). Given that \(x\) and \(y\) have symmetric roles, 
	we can assume that \(x \leqslant y\). If \(x > b\) then \(g(z) \neq g(x)\) as seen above. If \(y \leqslant b\) then 
	\(f(x) = g(x) = g(y) = f(y) \leqslant n\) and either \(z \leqslant a + b\) and \(f(z) \neq f(x)\) since \(f\) is a weakly 
	sun-free colouring or \(a + b + 1 \leqslant z \leqslant 2 b\) and \(g(z) = n + 1\); therefore \(g(z) \neq g(x)\). If 
	\(x \leqslant b\) and \(y > b\) then \(g(x) = f(x)\), \(g(y) = f(y - a)\) and \(g(z) = f(z - a)\).We have \(x \neq y - a\) 
	(otherwise, we would have \(a + b \geqslant z = 2 y  - a \geqslant 4 b + 4 - a > a + b\)) and thus \(f(x + y - a) \neq f(x)\),
	that is \(g(z) \neq g(x)\).
\item Colour \(n + 1\) verifies the additionnal constraints for the special colour. Indeed, \(b + 2 a \geqslant 4 b + 2\). 
	Hence, \(\forall (x, y) \in g^{-1}(\{n + 1\}), x + y \leqslant b + 2 a\).
\item Colours \(1, ..., n\) verify the additionnal constraints for the regular colours. Indeed, let 
	\((x, y) \in g^{-1}([\![1, n]\!])\) such that \(x + y > a + b\). Given that \(x\) and \(y\) have symmetric roles, 
	we can assume that \(x \leqslant y\). Necessarily \(y \geqslant 2 b + 2\). If \(x \leqslant b\) then 
	\(z = x + y \in [\![a + b + 1, a + 2 b]\!]\) and therefore \(\pi(z) \in [\![b + 1, 2 b]\!]\). Otherwise, 
	\(z = x + y \in [\![4 b + 4, 2 a + 2 b]\!]\) and therefore \(\pi(z) \in [\![b + 1, 2 b]\!]\). Hence, in both cases, 
	\(g(\pi(z)) \neq g(x)\).
\end{itemize}

@@

As in \hyperref[Corollary2.5]{Corollary 2.5}, the additive constant \hyperref[Theorem3.18]{Theorem 3.18} can be improved 
by weakening the hypotheses made on the last row. The principle behind it is the same as in \hyperref[Proposition2.6]{Proposition2.6}.

@@

Inequalities (\ref{WS(n+1)}) and (\ref{WS(n+2)}) were found by Rowley, they are detailed in \cite{RowleyWS}. 
Inequality (\ref{WS(n+3)}) is cannot be further improved and was found with a SAT solver. It uses the first 
sophistication explained in the \hyperref[ConstructionWS]{Subsection 3.3} in order to add
the last number in the first color. As for inequality (\ref{WS(n+4)}), it was obtained by combining a S-template with 
width 33 with a WS-template with width 4. The best template we could find with a computer search gives the 
inequality \(\WS(n+4) \geqslant 127 S(n) + 68\). It was also found with the SAT solver. In order to reduce the search space, 
we only looked for WS-templates of five colors which start with a near-optimal \(\WS(4)\) partition and we assumed that the 
special color was the last by order of appearance. We highly suspect that bettert WS-templates with \(n \geqslant 5\) 
colors can be found but one would have not to use the above assumptions. One may try to go over a different search space 
using a Monte-Carlo method, as in \cite{Bouzy2015AnAP}. This could be the suject of a future work.

>>>>

\begin{itemize}
\item Function \(g_{|[\![b + 1, a + b]\!]}\) is a sum-free colouring. Indeed, let \((x, y) \in 
	[\![b + 1, a + b]\!]^2\) such that \(g(x) = g(y)\). If \(g(x) = n + 1\) then \(z = x + y > 2 b + 1\) 
	and therefore, either \(z > a + b\) or \(f(z) \neq n + 1\). Otherwise, \(x + y > a + b\).
\item Function \(g\) is a weakly sum-free colouring. Indeed, let \((x, y) \in [\![1, a + b]\!]^2\) such that 
	\(z = x + y \leqslant a + b\) and \(g(x) = g(y)\). Given that \(x\) and \(y\) have symmetric roles, 
	we can assume that \(x \leqslant y\). If \(x > b\) then \(g(z) \neq g(x)\) as seen above. If \(y \leqslant b\) then 
	\(f(x) = g(x) = g(y) = f(y) \leqslant n\) and either \(z \leqslant a + b\) and \(f(z) \neq f(x)\) since \(f\) is a weakly 
	sum-free colouring or \(a + b + 1 \leqslant z \leqslant 2 b\) and \(g(z) = n + 1\); therefore \(g(z) \neq g(x)\). If 
	\(x \leqslant b\) and \(y > b\) then \(g(x) = f(x)\), \(g(y) = f(y - a)\) and \(g(z) = f(z - a)\).We have \(x \neq y - a\) 
	(otherwise, we would have \(a + b \geqslant z = 2 y  - a \geqslant 4 b + 4 - a > a + b\)) and thus \(f(x + y - a) \neq f(x)\),
	that is \(g(z) \neq g(x)\).
\item Colour \(n + 1\) verifies the additionnal constraints for the special colour. Indeed, \(b + 2 a \geqslant 4 b + 2\). 
	Hence, \(\forall (x, y) \in g^{-1}(\{n + 1\}), x + y \leqslant b + 2 a\).
\item Colours \(1, ..., n\) verify the additionnal constraints for the regular colours. Indeed, let 
	\((x, y) \in g^{-1}([\![1, n]\!])\) such that \(x + y > a + b\). Given that \(x\) and \(y\) have symmetric roles, 
	we can assume that \(x \leqslant y\). Necessarily \(y \geqslant 2 b + 2\). If \(x \leqslant b\) then 
	\(z = x + y \in [\![a + b + 1, a + 2 b]\!]\) and therefore \(\pi(z) \in [\![b + 1, 2 b]\!]\). Otherwise, 
	\(z = x + y \in [\![4 b + 4, 2 a + 2 b]\!]\) and therefore \(\pi(z) \in [\![b + 1, 2 b]\!]\). Hence, in both cases, 
	\(g(\pi(z)) \neq g(x)\).
\end{itemize}

@@

As in \hyperref[Corollary2.5]{Corollary 2.5}, the additive constant \hyperref[Theorem3.18]{Theorem 3.18} can be improved 
by weakening the hypotheses made on the last row. The principle behind it is the same as in \hyperref[Proposition2.6]{Proposition 2.6}.

@@

Inequalities (\ref{WS(n+1)}) and (\ref{WS(n+2)}) were found by Rowley, they are detailed in \cite{RowleyWS}. 
Inequality (\ref{WS(n+3)}) cannot be further improved and was found with a SAT solver. It uses the first 
sophistication explained in the \hyperref[ConstructionWS]{Subsection 3.3} in order to add
the last number in the first color. As for inequality (\ref{WS(n+4)}), it was obtained by combining a S-template with 
width 33 with a WS-template with width 4. The best template we could find with a computer search gives the 
inequality \(\WS(n+4) \geqslant 127 S(n) + 68\). It was also found with the SAT solver. In order to reduce the search space, 
we only looked for WS-templates of five colors which start with a near-optimal \(\WS(4)\) partition and we assumed that the 
special color was the last by order of appearance. We highly suspect that bettert WS-templates with \(n \geqslant 5\) 
colors can be found but one would have not to use the above assumptions. One may try to go over a different search space 
using a Monte-Carlo method, as in \cite{Bouzy2015AnAP}. This could be the suject of a future work.
